\documentclass{article}
\usepackage{hyperref}
\usepackage{listings}

\begin{document}

\title{Multi-agent movement behaviors in game AI}
\author{Jordan Lewis}
\maketitle

\begin{abstract}
A common problem in some games with large numbers of AI agents is encoding
individual behavior in such a way that a coherent group behavior is also
formed. For example, a squad of enemy troops might look more realistic moving
toward the player in a fixed (i.e. triangle) or fluid (i.e. follow a leader)
formation than simply all trying to run at the player directly. We will explore
various algorithms and methods for coordinated multi-agent movement in game
AI. Some of these behaviors will not require inter-agent communication of more
than just what's knowable from the physics environment, and some will - we will
also attempt to note the differences in efficacy for these two sorts of
algorithms.
\end{abstract}

\section {Introduction}
This paper attempts to explore a few different group AI behaviors. The paper
is accompanied by a demo that showcases these behaviors - read the README
in this directory for more information on how the demo works.

The behaviors we will examine hilight some different key decisions necessary
when designing a group AI: the choices between centralized and decentralized
information, and between uniform and nonuniform, behaviors across agents. I
define a uniform group AI behavior to be running identically across all agents,
as opposed to a non-uniform behavior which requires different algorithms to be
run on different agents, for leader behavior for example. I define a centralized
group AI algorithm as one that allows the agents to share state, as opposed to
a decentralized algorithm that doesn't allow inter-agent communicaton.
\section {Behaviors}
\subsection {Flocking}
Flocking is one of the most well-known group AI algorithms there is. It was
created by Craig Reynolds in 1987, for a demo called boids. It's a very simple
uniform and decentralized algorithm that produces realistic and interesting
results.

Flocking behavior is created by blending together three steering behaviors based
on the average behavior of the other agents in neighborhoods of differing size
around the current agent.
\begin{enumerate}
\item Separation ensures that agents don't get too close together, by steering
away from the average position of the agents in the neighborhood.
\item Cohesion ensures that agents clump together to form flocks, by steering
toward the average position of the agents in the neighborhood.
\item Matching ensures that agent clumps tend to move all together, by steering
to try to match the average heading and velocity of the agents in the
neighborhood.
\end {enumerate}

I used the values 5, 10, and 15 for separation, cohesion, and matching
respectively for my demo. This produced fairly good results, but typically the
clumps would tend to lose speed and reach a static equilibrium. I solved this
by adding a wander behavior with a low weight - this ensures that agents keep
moving around within their flock, which in turn motivates the flock to move as
a whole.

I implemented this algorithm with knowledge that the agents would not typically
be allowed to have - the actual positions, velocities, and orientations of all
the agents in the simulation. This is an optimization: perhaps the more
genuine approach would be to have each agent scan its neighborhood with raycasts
for other agents, and detect their velocity by sampling, but this is
very computationally intensive and would produce results that were at best as
interesting as the algorithm that uses global information.

\subsection {Follow the Leader}
Follow the leader is a very simple non-uniform, decentralized behavior. One
leader agent wanders around the world, and the other agents attempt to seek to
the leader's position.

Naively, this algorithm didn't produce very good results. The seek behavior
resulted in oscillatory behavior where the seekers would continually overshoot
the leader. This was because the wandering agent's velocity was never nearly
as high as the seeking agents', due to the random acceleration changes in
the wander behavior. Secondly, the agents would all ram into each other in a
mad dash to get to the leader's position.

The overshooting oscillation behavior was solved by using the arrive behavior
with a large slow-down radius around the leader. This produces smooth results
that appear as though the agents are trying to follow behind the leader at some
kind of respectful distance. The collision problem was solved by re-using the
separation steering algorithm from the flocking behavior.

\subsection {Static Formations}
Static formations are another kind of non-uniform but decentralized group AI
behavior. A static formation algorithm simply defines a relative position to
the leader given an integer spot. This can be used to make simple geometric
shapes for the agents to attempt to conform to.

I implemented two static formations in the demo: a line of agents abreast of
the leader, and a flying vee formation. These produce very nice looking results at minimal effort: with a simple inheritance relationship, an AI behavior writer
literally only has to specify a tiny 'getSpot' function. Here's the function I
wrote to define the line behavior.

\begin{lstlisting}
    pos = k.pos + k.vel.perp(Vec3f(0,1,0)).unit() * slot * (slot % 2 ? -1 : 1);
\end{lstlisting}
Given a kinematic k with position and velocity information of the leader, we
define a new position vector that's orthogonal to the leader's velocity vector,
alternating left and right, and scaled by the current agent's slot.

\subsection {Dynamic Formations}
Dynamic formations are similar to static formations in intent, but they offer
more flexibility in terms of managing lots of agents in one formation all at
onec. Instead of assigning a relative position to an agent based on only its
slot number, dynamic formations also take into account the number of total
agents participating in the system. This allows for more interesting resizing or
changing of the formation. I implemented a simple resizable circle formation
to demonstrate the dynamic resizability of the algorithm.

\section {Conclusion}


\section {Sources}
\begin{enumerate}
\item Individual-based models - Craig Reynolds - \url{www.red3d.com/cwr/ibm.html}
\item Steering behaviors - Craig Reynolds - \url{www.red3d.com/cwr/steer/}
\item Path planning algorithm of maintaining CGF team formation - Zheng et al. - \url{www.computer.org/portal/web/csdl/doi/10.1109/CSIE.2009.495}
\item Team AI: probabilistic pathfinding - John et al. - \url{portal.acm.org/citation.cfm?id=1234341.1234372&coll=ACM&dl=ACM&CFID=86476434&CFTOKEN=46450234}
\item Artificial Intelligence for Games - 2nd Ed - Millington and Funge
\end{enumerate}

\end{document}
